---
title: "Homework 8"
format: html
author: Priscilla Chen
editor: visual
embed-resources: true
code-fold: true
warning: false
---

### Exercise 1 (Supply and demand)

```{r}
#| include: false
library(reticulate)
library(dplyr)
library(tidyr)
library(readr)
library(vegabrite)
```

```{r}
data_url <- "https://calvin-data304.netlify.app/data/swd-lets-practice-ex-2-3.csv"
sales_data <- read_csv("https://calvin-data304.netlify.app/data/swd-lets-practice-ex-2-3.csv")

offsetbar <- vl_chart(width = 550, height = 400) |>
  vl_mark_bar() |>
  vl_fold(fold = c("demand", "capacity"), as = c("type", "value"))|>
  vl_encode_x("date:N", axis = list(labelAngle = 0), title = "date (from April 2019 to September 2020)", sort = c("January", "February", "March", "April", "May", "June", "July", 
                       "August", "September", "October", "November", "December")) |>
  vl_encode_y("value:Q", title = "Quantity") |>
  vl_encode_xOffset("type:N",sort = c("demand","capacity")) |>
  vl_encode_color("type:N", title = "")|>
  vl_encode_tooltip_array(list("demand","capacity","date"))|>
  vl_add_data(sales_data)
offsetbar
```

```{r}
layerbar <- vl_chart(width = 550, height = 400) |>
  vl_calculate(calculate = "datum.demand - datum.capacity", as = "unmet demand") |>
  vl_fold(fold = c("demand", "capacity", "unmet demand"), as = c("type", "value")) |>
  vl_filter(filter = ("datum.type === 'capacity' || datum.type === 'unmet demand'")) |>
  vl_mark_bar() |>
  vl_encode_x("date:T") |>
  vl_encode_y("value:Q", stack = "zero", title = "Quantity") |>
  vl_encode_size("type:N", sort = c("unmet demand", "capacity"), title = "") |>
  vl_encode_color("type:N", sort = c("unmet demand", "capacity"), title = "",legend = "") |>
  vl_scale_size(range = c(10,17))|>
  vl_encode_order("type:N", sort = c("capacity", "unmet demand")) |>
  vl_encode_tooltip_array(list("unmet demand","capacity"))|>
  vl_add_data(sales_data)
layerbar
```

```{r}
linechart2 <-vl_chart(width = 550, height = 400)|>
  vl_mark_line()|>
  vl_fold(fold = c("demand", "capacity"), as = c("type", "value"))|>
  vl_encode_x("date:T")|>
  vl_encode_y("value:Q", title = "Quantity")|>
  vl_encode_color("type:N", title = "")|>
  vl_add_data(sales_data)
linechart2
```

```{r}
unmetline <- vl_chart(width = 550, height = 400)|>
  vl_calculate(calculate = "datum.demand - datum.capacity", as = "unmet demand")|>
  vl_mark_line()|>
  vl_encode_x("date:T")|>
  vl_encode_y("unmet demand:Q", title = "Quantity unmet demand")|>
  vl_encode_tooltip_array(list("unmet demand","demand","capacity"))|>
  vl_add_data(sales_data)
unmetline
```

```{r}
stackedbar <- vl_chart(width = 550, height = 400) |>
  vl_calculate(calculate = "datum.demand - datum.capacity", as = "unmet demand") |>
  vl_fold(fold = c("demand", "capacity", "unmet demand"), as = c("type", "value")) |>
  vl_filter(filter = ("datum.type === 'capacity' || datum.type === 'unmet demand'")) |>
  vl_mark_bar() |>
  vl_encode_x("date:T") |>
  vl_encode_y("value:Q", stack = "zero", title = "Quantity") |>
  vl_encode_color("type:N", sort = c("capacity", "unmet demand"), title = "") |>
  vl_encode_order("type:N", sort = c("capacity", "unmet demand")) |>
  vl_encode_tooltip_array(list("unmet demand","capacity"))|>
  vl_add_data(sales_data)
stackedbar
```

```{r}
vertline <- vl_chart(width = 550, height = 400)|>
  vl_mark_line()|>
  vl_encode_x("date:T")|>
  vl_encode_y("demand:Q",title = "Quantity")|>
  vl_encode_y2("capacity:Q")|>
  vl_layer(vl_chart()|>
             vl_fold(fold = c("demand", "capacity"), as = c("type", "value"))|>
    vl_mark_point()|>
      vl_encode_x("date:T")|>
      vl_encode_y("value:Q")|>
      vl_encode_color(field = "type:N", title = "")|>
      vl_encode_fill(field = "type:N", title = ""))|>
  vl_encode_tooltip_array(list("demand","capacity"))|>
  vl_add_data(sales_data)
vertline
```

### Exercise 2 (Jobs)

a.  There was some question as to just what `perc` meant in this data set. Create a graphic that shows the sum of all the values of `perc` in each year separated by men and women. Explain what your graphic reveals about what `perc` means.

    ```{r}
    jobs_url <- 
      "https://cdn.jsdelivr.net/npm/vega-datasets@2.8.0/data/jobs.json"

    vl_chart(width = 500, height = 350)|>
      vl_mark_bar() |>
      vl_encode_x("year:Q")|>
      vl_scale_x(domainMin = 1850,domainMax = 2000)|>
      vl_encode_y("perc:Q")|>
      vl_scale_y(domainMax = 1)|>
      vl_encode_color("sex:N")|>
      vl_add_data_url(jobs_url)
    ```

-   perc is the number of (men or women) working at a particular job divided by the total number of people working in a particular year

b.  Now modify the graphic from class so that

    -   there are facets for 4 different years (your choice which years),

    -   scatterplots that show the percent of men and of women who worked in various jobs during those years, and

    -   `format=".0%"` is used (in the appropriate place) to display `perc` as a percentages (multiplied by 100 with a % sign appended).

    ```{r}
    vl_chart()|>
      vl_filter("datum.year === 1850 || datum.year === 1900 || datum.year === 1950 || datum.year === 2000")|>
      vl_mark_point()|>
      vl_pivot("sex", groupby = list("job", "year"), value = "perc") |>
      vl_encode_y("men:Q")|>
      vl_scale_y(type = "symlog", constant = 0.00001) |>
      vl_encode_x("women:Q")|>
      vl_scale_x(type = "symlog", constant = 0.00001) |>
      vl_facet("year:N", title = "") |>
      vl_resolve_axis_x("independent")|>
      vl_encode_tooltip_array(list("job", "men", "women")) |>
      vl_add_properties(title =  list( 
          text = "Percent of men verse women working various jobs",
          subtitle = "Each dot represents one occupation. Hover to see which occupation it is."
          ))|>
      vl_add_data_url(jobs_url)

    ```

c.  Redo part b but show the percentage of men who worked in various jobs and the percentage of women who worked in various jobs. (Part a should have convinced you that this is NOT what `perc` is in the original data set.)

    ```{r}
    library(jsonlite)
    jobs_data <- fromJSON(jobs_url)
    jobs_wide <- jobs_data |>
      # compute the total count per sex per year
      group_by(year, sex) |>
      mutate(total_count_sex = sum(count)) |>
      ungroup() |>
      # the percentage for each (sex, job) combination
      mutate(perc_sex = count / total_count_sex) 


    ```

    ```{r}
    vl_chart()|>
      vl_filter("datum.year === 1850 || datum.year === 1900 || datum.year === 1950 || datum.year === 2000")|>
      vl_mark_point()|>
      vl_pivot("sex", groupby = list("job", "year"), value = "perc_sex") |>
      vl_encode_y("men:Q")|>
      vl_scale_y(type = "symlog", constant = 0.00001) |>
      vl_encode_x("women:Q")|>
      vl_scale_x(type = "symlog", constant = 0.00001) |>
      vl_facet("year:N", title = "") |>
      vl_resolve_axis_x("independent")|>
      vl_encode_tooltip_array(list("job", "men", "women")) |>
      vl_add_properties(title =  list( 
          text = "Percent of men verse women working various jobs",
          subtitle = ""
          ))|>
      vl_add_data(jobs_wide)
    ```

### Exercise 3 (Price at the pump)

Create graphic(s) to compare gas prices in US \$ per gallon. Take advantage of the countries/regions/codes data to do things like use 3-letter codes for countries, group them by region, etc.

Do any countries fail to get matched?

-   YES. 28 countries in the country in the gas prices dataset cannot find a match in the country dataset.

#### Genderal Trend of Gas Prices

```{r}
gas_data <- read_csv("https://calvin-data304.netlify.app/data/pump_price_for_gasoline_us_per_liter.csv")
country_data <- read_csv("https://raw.githubusercontent.com/lukes/ISO-3166-Countries-with-Regional-Codes/master/all/all.csv")

# Left join
joined_data <- gas_data |>
  left_join(country_data, by = c("country" = "name"))

# check how many countries are not matched.
na_count <- sum(is.na(joined_data$`alpha-3`))
# na_count
```

```{r}
years <- as.character(1991:2016)

vl_chart(width = 550, height = 350) |>
  vl_fold(fold = years, as = c("year", "gas_price")) |>
  vl_calculate(calculate = "random()%3", as="random") |>
  vl_mark_point(filled = TRUE) |>
  vl_encode_x("year:O",axis = list(labelAngle = 0)) |>
  vl_encode_xOffset("random")|>
  vl_encode_y("gas_price:Q", title = "gas prices in US $ per gallon") |>
  vl_encode_color("year", title = "", legend = "")|>
  vl_resolve_axis_x("independent")|>
  vl_encode_tooltip_array(list("year","country","gas_price")) |>
    vl_add_properties(title =  list( 
          text = "General Trend of Gas Prices across Years",
          subtitle = "Each point represent the gas price of a particular country in a particular. Hover to see details"
          ))|>
  vl_add_data(joined_data)
```

#### Compare general trend of gas prices between regions

```{r}
years <- as.character(1991:2016)
# continents ordered by who has the most countries
order <- c("Africa","Asia","Europe","Americas", "Oceania", "null")

vl_chart(width = 450, height = 350) |>
  vl_fold(fold = years, as = c("year", "gas_price")) |>
  vl_calculate(calculate = "random()%3", as="random") |>
  vl_mark_point() |>
  vl_encode_x("year:O",axis = list(labelAngle = 0)) |>
  vl_encode_y("gas_price:Q", title = "gas prices in US $ per gallon") |>
  vl_encode_xOffset("random")|>
  vl_encode_color("year", title = "", legend = "")|>
  vl_facet("region:N", title = "", sort = order)|>
  vl_resolve_axis_x("independent")|>
  vl_encode_tooltip_array(list("year","country","gas_price")) |>
    vl_add_properties(title =  list( 
          text = "Compare genderal trend of gas prices between regions",
          subtitle = "Each point represent the gas price of a particular country in a particular. Hover to see details"
          ))|>
  vl_add_data(joined_data)
```

#### compare regions in years

```{r}
years <- 1991:2016
# continents ordered by who has the most countries
order <- c("Africa","Asia","Europe","Americas", "Oceania", "null")

vl_chart(width = 450, height = 350) |>
  vl_fold(fold = as.character(years), as = c("year", "gas_price")) |>
  vl_calculate(calculate = "random()%2 -1", as="random") |>
  vl_add_parameter("year", value = 1991,
    bind = list(input = "range",min = 1991, max = 2016, step = 1)
  ) |>
  vl_filter("toNumber(datum.year) == year") |>
  vl_mark_point() |>
  vl_encode_x("region:N",axis = list(labelAngle = 0), sort = order) |>
  vl_encode_y("gas_price:Q", title = "gas prices in US $ per gallon") |>
  vl_encode_xOffset("random")|>
  vl_encode_color("region:N", title = "",legend = "", sort = order)|>
  vl_encode_tooltip_array(list("year","country","gas_price")) |>
      vl_add_properties(title =  list( 
          text = "Compare gas prices between regions in a year",
          subtitle = "Each point represent the gas price of a particular country in a particular. Hover to see details."
          ))|>
  vl_add_data(joined_data)
```
