---
title: "Homework 9"
format: html
editor: visual
embed-resources: true
code-fold: true
---

```{r}
#| include: false
library(dplyr)
library(altair)
library(jsonlite)
library(mosaic)
library(reticulate)
library(readr)
library(vegabrite)
```

```{python}
#| include: false
import pandas as pd
import altair as alt
```

**Exercise 1 (Wilke)**

a.  Read chapter 4 of of Wilke, List two or three things you learned about using color from this chapter.

-   "Multi-hue scales tend to follow color gradients that can be seen in the natural world, such as dark red, green, or blue to light yellow, or dark purple to light green." This is interesting. I've wondered how people construct muti-hue scales, and this explaination makes sense.

-   Color as a mean to highlight is an interesting use I don't often think about. The example with taxes and Louisiana is a great example of highlighting the particular story the graph is trying to convey.

b.  Read chapter 5 of of Wilke. Chapter 5 list a wide range of graphic types. What is the organization scheme Wilke uses to group them? What is the most interesting thing you learned from this chapter?

-   Wilke grouped them by what type of story are they trying to convey ("Amounts", "Distribution", "Porportion", "X-Y relationship", "Geospacial", and "Uncertainty")

-   I was intrigued by the density contour graph and the correlogram graph in the x-y relationship section. These are very interesting ways that can convey more than 2 quantities.

c.  By now, you should be able to make quite a few of the graphics shown in chapter 5. Are there some you would not know how to make (yet) based on what we have done in class? List several that you don’t know how to make – choose the ones you would be most interested in learning how to make and say why you are interested.

-   Voilin plot: Voilin plots can not only show some descriptive statistics like the box plot, but can also visualize the shape of the distribution.\
-   Density Contours: It is an interesting way to visualize where the data is concentrated, where the clusters of data are at.
-   Mosaic Plot: It's an interesting and efficient way to visualize comparisons between categorical variables.

d.  See if you can figure out how to make at least one of the types of graphic listed in part c.  You may need to do some experimenting or to consult the documentation or example galleries (for Vega-Lite, vegabrite, altiar, Altair, etc.).

```{python}
Chem_admit = pd.read_csv("/rprojects/jc253/DATA304/docs/ChemAdmission.csv")
alt.Chart(Chem_admit, width=100).transform_density(
    'High School Transcript GPA',
    as_=['High School Transcript GPA', 'density'],
    extent=[2, 6],
    groupby=['Gender']
).mark_area(orient='horizontal').encode(
    alt.X('density:Q')
        .stack('center')
        .impute(None)
        .title(None)
        .axis(labels=False, values=[0], grid=False, ticks=True),
    alt.Y('High School Transcript GPA:Q'),
    alt.Color('Gender:N'),
    alt.Column('Gender:N')
        .spacing(0)
        .header(titleOrient='bottom', labelOrient='bottom', labelPadding=0)
).configure_view(
    stroke=None
)
```

e.  Chapters 6 - 16 are organized by graphic types. You have already read some of this material. Pick one chapter that is related to your graphic in part d and give some examples of how what you learned in that chapter affected how you created your graphic.

-   Chapter 9 talked about voilin plot.
-   I started making my voilin plot through grouping by Program Interest, and I found that while some data for program interest worked well, others had a weird shape (a vertical line). The book helped me think about this: "Before using violins to visualize distributions, verify that you have sufficiently many data points in each group to justify showing the point densities as smooth lines". I think those weird shapes might be because there's too few data values for that program interest.

**Exercise 2 (More pie)**

```{python}
survey_response = pd.read_csv("https://calvin-data304.netlify.app/data/likert-survey.csv")
```

a.  Create side-by-side pie charts with good labeling and appropriate choice of colors. Some more things to consider: Should you present using counts or percentages? How will you deal with the nonresponders?

```{python}
this_year = survey_response[survey_response["year"] == "this year"]
last_year = survey_response[survey_response["year"] == "last year"]


base = alt.Chart(this_year, width = 300, height = 300).mark_arc().encode(
  theta = alt.Theta("count:Q").stack(True), 
  color = alt.Color("number:O", legend = None),
  order = alt.Order("number", sort = "descending")
)
pie_1 = base.mark_arc(outerRadius=120)

text_chart = base.mark_text(radius = 80, size = 10).encode(
  text = "count:Q",
  color = alt.value("white"),
  order = alt.Order("number", sort = "descending")
)
text_chart2 = base.mark_text(radius = 160, size = 10).encode(
  text = "response:N",
  color = alt.value("black"),
  order = alt.Order("number", sort = "descending")
)
Total_chart_1 = pie_1 + text_chart + text_chart2
# Total_chart_1

base_2 = alt.Chart(last_year, width = 300, height = 300).mark_arc().encode(
  theta = alt.Theta("count:Q").stack(True), 
  color = alt.Color("number:O", legend = None),
  order = alt.Order("number", sort = "descending")
)
pie_2 = base_2.mark_arc(outerRadius=120)

text_chart = base_2.mark_text(radius = 80, size = 10).encode(
  text = "count:Q",
  color = alt.value("white"),
  order = alt.Order("number", sort = "descending")
)
text_chart2 = base_2.mark_text(radius = 150, size = 10).encode(
  text = "response:N",
  color = alt.value("black"),
  order = alt.Order("number", sort = "descending")
)
Total_chart_2 = pie_2 + text_chart + text_chart2
# Total_chart_2

Two_year = alt.hconcat(Total_chart_1,Total_chart_2)
Two_year
```

For the pie graphs, I used counts, because the pie chart already shows the proportion of a response as a whole. I treated the "no response" as one response because it still represent a proportion of people who didn't respond to the question. It can be a little misleading and less comprehensive if their responses are taken out.

b.  Create some other kind of graphic to visualize the same data.

**Point Chart**

```{python}
Order = ["strongly agree", "agree", "neither agree nor disagree","disagree", "strongly disagree", "no response"]
survey = alt.Chart(
  survey_response,width = 300, height = 300
  ).mark_point(
).encode(
  x = alt.X("count:Q"),
  y = alt.Y("response:N", title = "",sort=Order),
  color = alt.Color("year:N", title = ""),
  fill = alt.Fill("year:N", title = "")
)
survey
```

**Bar Chart**

```{python}
survey_wide = survey_response.pivot(index="response", columns="year", values="count").reset_index()
survey_wide["change"] = survey_wide["this year"]- survey_wide["last year"]
survey_wide["direction"] = (survey_wide["change"] >= 0).astype(bool)
survey_bar = alt.Chart(survey_wide, width = 300, height = 300
).mark_bar().encode(
  x = alt.X("change:Q", title = "Change in Counts of Response from Last Year to This Year"),
  y = alt.Y("response:N", title = "", sort = Order),
  color = alt.Color("direction:N", title = "",legend=None)
)
survey_bar
```

c.  Comment on the pros and cons of your two (or more) graphics. Which would you recommend (for what purposes)?

-   The side-by-side pie chart is better at presenting and comparing the proportion of a particular response with another response within one year. It is not great at comparing the responses between last year and this year. In addition, if there is response other than the 6 listed in the dataset, the pie graphics might be misleading in suggesting that these 6 responses make up the whole pool of survey responses.
-   The point chart is better at comparing the value of each response between years as well as with other responses. It is also more space efficient. However, this point chart does not convey that those 6 responses make up all the survey responses as a whole intuitively.
-   The bar chart is better at viewing the change from last year to this year, especially, it could communicate the direction and magnitude of the change visually. However, it does not show the count of each responses.

**Exercise 3 (Create a choropleth map)**

```{r}
vega_data <- import_vega_data()
us_map_url <- vega_data$us_10m$url
state_map <-
  vl_chart() |>
  vl_add_data(
    url = us_map_url, 
    format = list(type = "topojson", feature = "states")) |>
  vl_add_properties(projection = list(type = "albersUsa")) |>
  vl_mark_geoshape(fill = "transparent", stroke = "steelblue")|> vl_add_properties(width = 350, height = 200)
```

```{r}
obesity_colored <- state_map |>
  vl_lookup(
    lookup = "id",                  
    from = 
      list(
        data = list(url = vega_data$obesity$url), 
        key = "id",                
        fields = list("rate"))) |>  
  vl_encode_stroke(value = "transparent") |>
  vl_encode_fill("rate:Q") |>
  vl_legend_fill(title = "Obesity rate", format = ".0%", type = "gradient") |>
  vl_add_properties(width = 350, height = 200)

obesity_colored
```
